---
title: "MATH 3070 Lab Project 1"
author: "Patrick Ekel"
date: "17 May 2017"
output:
  html_document:
    toc: TRUE
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 1.1)
*Use R as you would a calculator to find numeric answers to the following:*

** '*' is used to explicitly define multipliation **
** eval function is used to evaluate expression **


1. $1 + 2(3 + 4)$

```{r}
# '*' was added to explicity define multiplication
eval(1+2*(3+4))
```

2. $4^3 + 3^{2 + 1}$

```{r}
# '^' is used to signify exponent
eval(4^3+3^(2+1))
```


3. $\sqrt{(4 + 3)(2 + 1)}$

```{r}
# sqrt function is used to find square root
eval(sqrt((4+3)*(2+1)))
```


4. $\left(\frac{1+2}{3 + 4} \right)^2$

```{r}
# parenthesss are crucial for order of ops
eval(((1+2)/(3 + 4))^2)
```

## Problem 2 (Verzani problem 1.4)
*Use R to compute the following:*


$$\frac{0.25 - 0.2}{\sqrt{0.2(1-0.2)/100}}$$

```{r}
# sqrt function is used for square root
eval((0.25 - 0.2)/sqrt(0.2*(1-0.2)/100))
```


## Problem 3 (Verzani problem 1.7)
*The `exec.pay` (**UsingR**) data set is available after loading the package **UsingR**. Load the package, and inspect the data set. Scan the values to find the largest one.*

```{r}
# mac function is used to find max data of argument exec.pay
# exec.pay is the pay variable o f the ceo data set
library(UsingR)
max(exec.pay)
```

