---
title: "MATH 3070 Lab Project 2"
author: "Your Name"
date: "24 May 2017"
output:
  html_document:
    toc: TRUE
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***


## Problem 1 (Verzani problem 1.8)
*For the `exec.pay` data set, apply the functions `mean()`, `min()`, and `max()`. What are the values found?*

```{r}
# min, max and mean of exec.pay stored in a vector
MinMaxMean <- c(min(exec.pay),max(exec.pay),mean(exec.pay))
MinMaxMean # show min, max, mean values
```

## Problem 2 (Verzani problem 1.10)
*The `Orange` data set is stored as a data frame with three variables. What are the three variables?*

```{r}
Orange # print entire data frame
names(Orange) # print variables (columns) of data
```

## Problem 3 (Verzani problem 2.1)
*Look up the documentation of the `length()` function. Then create a vector named `p` using `c()` with values `2 3 5 7 11 13 17 19` and use `length()` to check its length.*

```{r}
?length # review documentation on length() function
p <- c(2,3,5,7,11,13,17,19) # instantiate p to a vector
length(p) # print length of p
```

## Problem 4 (Verzani problem 2.4)
*Create the following sequences, using `:`, `seq()`, or `rep()` as appropriate:*

1. `"a" "a" "a" "a" "a" "a"`

```{r}
# Your solution here
rep("a", 6)
```

2. `1 3 ... 99` *(the odd numbers)*

```{r}
# Your solution here
seq(1, 99, 2)
```

3. `1 1 1 2 2 2 3 3 3`

```{r}
# Your solution here
j<-c(rep(1, 3), rep(2, 3), rep(3, 3))################################################# CHANGE THISSSSSSS    ###################################################

```

## Problem 5 (Verzani problem 2.9)
*An experiment had 10 different trials. Create a character vector with 100 different names for the trials, e.g., `"Trial 1" "Trial 2" ...`*

```{r}
# Your solution here
#char_vec <- 

```


## Problem 6 (Verzani problem 2.3)
*Let our small data set be* `2 5 4 10 8`.

1. *Enter this data into a data vector `x`.*

```{r}
# Your code here
```

2. *Find the square of each number.*

```{r}
# Your code here
```

3. *Subtract 6 from each number.*

```{r}
# Your code here
```

4. *Subtract 9 from each number and then square the answer.*

```{r}
# Your code here
```

## Problem 7 (Verzani problem 2.15)
*The negation operator `!` is used to reverse Boolean values. For example:*

```{r}
A <- c(TRUE, FALSE, TRUE, TRUE)
!A
```

*One of De Morgan's laws in R code is `!(A & B) == !A | !B`. Verify this with `B <- c(FALSE, TRUE, FALSE, TRUE)` and `A` as above.*

```{r}
# Your code here
```

## Problem 8 (Verzani problem 2.19)
*You track your commute times for two weeks (ten days), recording the following times in minutes:*

`17 16 20 24 22 15 21 15 17 22`

*Enter these into R. Use the function `max()` to find the longest commute time, the function `mean()` to find the average, and the function `min()` to find the minimum.*

```{r}
# Your code here
```

*Oops, the 24 was a mistake. It should have been 18. How can you fix this? Do so, and then find the new average.*

```{r}
# Your code here
```

*How many times was your commute 20 minutes or more? What percentage of your commutes are less than 18 minutes long?*

```{r}
# Your code here
```

## Problem 9 (Verzani problem 2.25)
*Write a function `isprime()` that checks if a number $x$ is prime by dividing $x$ by all the values in $2, ..., x - 1$ then checking to see if there is a remainder of 0. The expression `a %% b` returns the remainder of `a` divided by `b`.*

```{r}
# Your code here
```

***BONUS:** You do not need to check all numbers from 2 to $x - 1$ to see if a number is prime. What is the largest you would need to go for an arbitrary $x$? Create a new function, `isprime2()`, that implements this better (yet still slow) method.*

```{r}
# Your code here
```

## Problem 10 (Verzani problem 2.31)
*For the data sets `bumpers` (**UsingR**), `firstchi` (**UsingR**), and `math` (**UsingR**), make histograms. Try to predict the mean, median, and standard deviation from the graphic. Check your guesses with the appropriate R commands.*

```{r}
# Your code here
```

## Problem 11 (Verzani problem 2.32)
*Fit a density estimate to the data set `pi2000` (**UsingR**). Compare with the appropriate histogram. Why might you want to add an argument like `breaks = 0:10-.5` to `hist()`? (Hint: read the documentation of `hist()` to see what setting this argument does.)*

```{r}
# Your code here
```

## Problem 12 (Verzani problem 2.34)
*The data set `DDT` (**MASS**) contains independent measurements of the pesticide DDT on kale. Make a histogram and a boxplot of the data. From these, estimate the mean and standard deviation. Check your answers with the appropriate functions.*

```{r}
# Your code here
```

## Problem 13 (Verzani problem 2.43)
*The `time` variable in the `nym.2002` data set contains the time to finish the 2002 New York City Marathon for a random sample of the finishers.*

1. *What percent ran the race in under 3 hours?*

```{r}
# Your code here
```

2. *What is the time cutoff for the top 10%? The top 25%?*

```{r}
# Your code here
```

3. *What time cuts off the bottom 10%?*

```{r}
# Your code here
```

## Problem 14 (Verzani problem 2.63)
*The `lorem` variable in UsingR contains 5 paragraphs of dummy typesetting text that has been in use for centuries. What is the most common letter used? To answer this, you can break a character value into letters by the idiom `unlist(strsplit(x,""))` where `x` is character data.*

```{r}
# Your code here
```

*Make a bar chart to visualize the distribution of characters. What do you notice?*

```{r}
# Your code here
```

## Problem 15 (Verzani problem 4.1)
*The data set `UScereal` (**MASS**) contains data on cereals sold in the United States in 1993. For this data set, answer the following questions:*

1. *How many rows does the data frame have? Columns?*

```{r}
# Your code here
```

2. *How many different manufacturers are included?*

```{r}
# Your code here
```

3. *How many vitamin categories are included?*

```{r}
# Your code here
```

4. *How many cereals have a sugar level above 10?*

```{r}
# Your code here
```

5. *What is the mean calorie value for cereals with more than 5 grams of fat? Less than or equal to 5?*

```{r}
# Your code here
```

6. *What is the mean calorie value for cereals on the middle shelf (2)?*

```{r}
# Your code here
```

## Problem 16 (Verzani problem 4.2)
*R uses lists for many purposes behind the scenes. For example, the output of `lm(mpg ~ wt, data=mtcars)` returns a list. Create this object, then answer the following:*

1. *How many components does this list have?*

```{r}
# Your code here
```

2. *What are the names of the components?*

```{r}
# Your code here
```

3. *What kind of data is held in the `residuals` variable?*

```{r}
# Your code here
```