---
title: "MATH 3070 Lab Project 4"
author: "Patrick Ekel"
date: "7 June 2017"
output:
  html_document:
    toc: TRUE
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

```{r, tidy=TRUE}
library(UsingR)
library(ggplot2)
library(lattice)
#
#
#
#
#
###     as.factor(cyl)
#
#
#
#
#
```

## Problem 1 (Verzani problem 5.1)
*For the `UScereal` (**MASS**) data set, create a scatter plot of `calories` modeled by `sugars` using the `shelf` variable to create different plot characters. Add a legend to indicate the shelf number. Is there any patterns? (Use base R plotting for this problem.)*

```{r, tidy=TRUE}
# Your code here
plot(UScereal$calories, UScereal$sugars)
with(UScereal, plot(calories, sugars, col = as.numeric(shelf), pch = as.numeric(shelf), xlab = "calories", ylab = "sugars", main = "Cereal Calories vs Sugars by Shelf") )
legend(397.5, 21.5, c("shelf1", "shelf2", "shelf3"), col = 1:3, pch = 1:3)
```

## Problem 2 (Verzani problem 5.4)
*For the data set `batting` (**UsingR**) make a bubble plot of home runs hit (`HR`) modeled by hits (`H`) where the scale factor for each point is given by `sqrt(SO)/10`. Is there any story to be told by the size of the points? (You must use base R plotting for this problem.)*

```{r, tidy=TRUE}
# bubble plot of batting H, HR data where bubble size is set by cex parameter
plot(HR ~ H, data = batting, pch = 16, col=rgb(red = 0, green = 0, blue = 0, alpha = .25), cex = sqrt(SO)/5)
# The general trend is as the amount of home runs and hits increase, so do the amount of strike outs, as # illustraded by the increading size of the bubbles
```

## Problem 3 (Verzani problem 5.6)
*For the `batting` (**UsingR**) data set, make parallel boxplots of the batting average (`H/AB`) for each team. Which team had the greatest median average? (Use **lattice** functions for this problem.)*

```{r, tidy=TRUE}
# boxplot of batting average (H/AB) for each different team 
bwplot(teamID ~ H/AB, data = batting)
  # ANA had the greatest median average

```


## Problem 4 (Verzani problem 5.7)
*For the `mtcars` data set, produce graphics of the following using **ggplot2**:*

1. *Boxplots for miles per gallon (`mpg`) for groups defined by the number of gears (`gear`).*

```{r, tidy=TRUE}
# ggplot2 uses qplot function name. geom specifies the type of plot while group specifies the critera to group by
qplot(gear, mpg, data = mtcars, geom = "boxplot", group = gear)
```

3. *A scatterplot of `mpg` modeled by horsepower (`hp`). Create facets by the number of cylinders (`cyl`) and `gear`.*

```{r, tidy=TRUE}
# Again using ggplot2 
# now using default ploit which is scatter but adding a two facets to the grid structure->cyl and gear
p<-qplot(hp, mpg, data = mtcars, color = hp) + facet_grid(cyl~gear)
print(p)
```

## Problem 5
*Use any plotting system (base R, **lattice**, **ggplot2**) to create plot involving at least three variables in the `batting` (**UsingR**) data set, showing a relationship not yet visualized in the lecture, the textbook, or this assignment. Explain the relationship you explored and any interesting findings. **Bonus points will be given for plots that I consider exceptionally clean, clear, and insightful, accompanied with good analyses of what you found.***

```{r, tidy = TRUE}
# Your code here
#attach(batting)
str(batting)
head(batting)
library(ggplot2)

batting$lgID<-factor(batting$lgID, levels=c("AL", "NL"), labels=c("american", "national"))

runs_stolen<-qplot(SB, R, data=batting, geom=c("point", "smooth"), method="lm", formula=y~x, color=lgID, main = "Regression of Runs ~ Stolen Bases", xlab="Stolen Bases", ylab="Runs Scored")



qplot(HR, R, data=batting, geom=c("point", "smooth"), method="lm", formula=y~x, color=lgID, main="balls")



# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}



library(ggplot2)
li

r_dub<-ggplot(batting, aes(x=DOUBLE, y = R, color=lgID)) + geom_point(alpha=.2) + geom_smooth(alpha=.3, size=1) + ggtitle("R~Doubles") + xlab("Doubles") + ylab("Runs")
r_trip<-ggplot(batting, aes(x=TRIPLE, y = R, color=lgID)) + geom_point(alpha=.2) + geom_smooth(alpha=.3, size=1) + ggtitle("R~Triples") + xlab("Triples") + ylab("Runs")
r_hr<-ggplot(batting, aes(x=HR, y = R, color=lgID)) + geom_point(alpha=.2) + geom_smooth(alpha=.3, size=1) + ggtitle("R~Home Runs") + xlab("Home Runs") + ylab("Runs")
r_stolen<-ggplot(batting, aes(x=SB, y=R, color=lgID)) + geom_point(alpha=.2) + geom_smooth(alpha=.3, size=1) + ggtitle("R~Stolen Bases") + xlab("Stolen Bases") + ylab("Runs")

multiplot(r_dub, r_trip, r_hr, r_stolen, cols =2)




```
