---
title: "MATH 3070 Lab Project 4"
author: "Your Name"
date: "7 June 2017"
output:
  html_document:
    toc: TRUE
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

```{r, tidy=TRUE}
library(UsingR)
library(ggplot2)
library(lattice)
#
#
#
#
#
###     as.factor(cyl)
#
#
#
#
#
```

## Problem 1 (Verzani problem 5.1)
*For the `UScereal` (**MASS**) data set, create a scatter plot of `calories` modeled by `sugars` using the `shelf` variable to create different plot characters. Add a legend to indicate the shelf number. Is there any patterns? (Use base R plotting for this problem.)*

```{r, tidy=TRUE}
# Your code here
plot(UScereal$calories, UScereal$sugars)
with(UScereal, plot(calories, sugars, col = as.numeric(shelf), pch = as.numeric(shelf), xlab = "calories", ylab = "sugars", main = "Cereal Calories vs Sugars by Shelf") )
legend(397.5, 21.5, c("shelf1", "shelf2", "shelf3"), col = 1:3, pch = 1:3)
```

## Problem 2 (Verzani problem 5.4)
*For the data set `batting` (**UsingR**) make a bubble plot of home runs hit (`HR`) modeled by hits (`H`) where the scale factor for each point is given by `sqrt(SO)/10`. Is there any story to be told by the size of the points? (You must use base R plotting for this problem.)*

```{r, tidy=TRUE}
# bubble plot of batting H, HR data where bubble size is set by cex parameter
plot(HR ~ H, data = batting, pch = 16, col=rgb(red = 0, green = 0, blue = 0, alpha = .25), cex = sqrt(SO)/5)
# The general trend is as the amount of home runs and hits increase, so do the amount of strike outs, as illustraded by the increading size of the bubbles


#
#            ask about this
#
```

## Problem 3 (Verzani problem 5.6)
*For the `batting` (**UsingR**) data set, make parallel boxplots of the batting average (`H/AB`) for each team. Which team had the greatest median average? (Use **lattice** functions for this problem.)*

```{r, tidy=TRUE}
# so this is all using base R...not lattice..
length_range <- range(batting$H/batting$AB)
# Manually split into three datasets
batting_ANA <- subset(batting, subset = teamID == "ANA")
batting_ARI <- subset(batting, subset = teamID == "ARI")
batting_DET <- subset(batting, subset = teamID == "DET")
batting_OAK <- subset(batting, subset = teamID == "OAK")
batting_TBA <- subset(batting, subset = teamID == "TBA")
batting_MIL <- subset(batting, subset = teamID == "MIL")
# save old par
old_par <- par()
# change the settings
par(mfrow = c(1,6))
#boxplot(batting_ANA$H/batting_ANA$AB, data = batting_ANA, ylim = length_range, main = "ANA")
#boxplot(batting_ARI$H/batting_ARI$AB, data = batting_ARI, ylim = length_range, main = "ARI")
#boxplot(batting_DET$H/batting_DET$AB, data = batting_DET, ylim = length_range, main = "DET")
#boxplot(batting_OAK$H/batting_OAK$AB, data = batting_OAK, ylim = length_range, main = "OAK")
#boxplot(batting_TBA$H/batting_TBA$AB, data = batting_TBA, ylim = length_range, main = "TBA")
#boxplot(batting_MIL$H/batting_MIL$AB, data = batting_MIL, ylim = length_range, main = "MIL")

head(batting_ANA)
boxplot(H/AB, data = batting_ANA, ylim = length_range, main = "ANA")
boxplot(H/AB, data = batting_ARI, ylim = length_range, main = "ARI")
boxplot(H/AB, data = batting_DET, ylim = length_range, main = "DET")
boxplot(H/AB, data = batting_OAK, ylim = length_range, main = "OAK")
boxplot(H/AB, data = batting_TBA, ylim = length_range, main = "TBA")
boxplot(H/AB, data = batting_MIL, ylim = length_range, main = "MIL")


# reset par to its original settings once done
#par(old_par)
#
#
#
#   this didnt work for me if not using batting_ANA$H/batt.....need to be that specififc...
#
#
#   ****************?????????????????Lattice???????????
#
#
```

```{r, tidy=TRUE}

bwplot(batting_ANA$H/batting_ANA$AB, data = batting_ANA, ylim = length_range, main = "ANA")
bwplot(batting_ARI$H/batting_ARI$AB, data = batting_ARI, ylim = length_range, main = "ARI")
bwplot(batting_DET$H/batting_DET$AB, data = batting_DET, ylim = length_range, main = "DET")
bwplot(batting_OAK$H/batting_OAK$AB, data = batting_OAK, ylim = length_range, main = "OAK")
bwplot(batting_TBA$H/batting_TBA$AB, data = batting_TBA, ylim = length_range, main = "TBA")
bwplot(batting_MIL$H/batting_MIL$AB, data = batting_MIL, ylim = length_range, main = "MIL")

```


## Problem 4 (Verzani problem 5.7)
*For the `mtcars` data set, produce graphics of the following using **ggplot2**:*

1. *Boxplots for miles per gallon (`mpg`) for groups defined by the number of gears (`gear`).*

```{r, tidy=TRUE}
# Your code here
mtcars
iris
qplot(gear, mpg, data = mtcars, geom = "boxplot")
qplot(Species, Petal.Length, data = iris, geom = "boxplot")



#
#
#
#   why so much trouble graphing seperatly(based on gear....)
#
#

```

3. *A scatterplot of `mpg` modeled by horsepower (`hp`). Create facets by the number of cylinders (`cyl`) and `gear`.*

```{r, tidy=TRUE}
# Your code here
mtcars
p<-qplot(hp, mpg, data = mtcars, color = hp) + facet_grid(cyl~gear)
print(p)
```

## Problem 5
*Use any plotting system (base R, **lattice**, **ggplot2**) to create plot involving at least three variables in the `batting` (**UsingR**) data set, showing a relationship not yet visualized in the lecture, the textbook, or this assignment. Explain the relationship you explored and any interesting findings. **Bonus points will be given for plots that I consider exceptionally clean, clear, and insightful, accompanied with good analyses of what you found.***

```{r, tidy = TRUE}
# Your code here

```