---
title: "MATH 3070 Lab Project 6"
author: "Patrick Ekel"
date: "21 June 2017"
output:
  html_document:
    toc: TRUE
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

*Because randomization is used in this assignment, I set the seed here, in addition to beginning each code block. **Do not change the seed!***

```{r}
set.seed(6222016)
```

## Problem 1 (Verzani problem 7.1)
*Simulate 1000 rolls of a pair of dice, and compute the sum of each pair. Which is more common, a roll of 7 or 8?*

```{r, echo=FALSE}
# Seed reset; DO NOT CHANGE THE SEED!
set.seed(6222016)
```

```{r, tidy=TRUE}
# Store a replated random number generation 1000 times with values 1-6
# do this twice to replicate two die
dice<-replicate(2, sample(6, 1000, replace=T))
tail(dice)
# sum the amount of times the sum of dice 1 and 2 is 7
sum(dice[,1]+dice[,2] == 7) # 167
sum(dice[,1]+dice[,2] == 8) # 149
# Roling a 7 is more common than rolling an 8
```

## Problem 2 (Verzani problem 7.2)
*For the `rivers` data set, take 1000 random samples of size 10 Compare the mean of the sample means computed from these samples, with the sample mean of the data in `rivers`.*

```{r, echo=FALSE}
# Seed reset; DO NOT CHANGE THE SEED!
set.seed(6222016)
```

```{r, tidy=TRUE}
# Calculate sample mean
samp_mean<-replicate(1000, mean(rnorm(10, mean =mean(rivers))))
mean(samp_mean)
# The mean of the sample data and the mean of the actual data is very very similar. THis is becasue of the large sample size
mean(rivers)
```

## Problem 3
*The data set `Melanoma` (**MASS**) includes data for 205 Danish patients with malignant melanoma. The variable `time` describes survival time in days, and `sex` describes the sex of the patient. Does survival time differ between the sexes?*

1. *Find $\bar{x}_{\text{men}} - \bar{x}_{\text{women}}$, the mean difference in survival time (`time`) between men and women (`sex`) in `Melanoma`.*

```{r, tidy = TRUE}
# subset the time data into female and male data
library(MASS)
time_male<-subset(Melanoma$time, Melanoma$sex==1) # 1 = male
time_female<-subset(Melanoma$time, Melanoma$sex==0) # 0 = female
# sample male and female data, caluclate mean
samp_mean_male<-replicate(1000, mean(rnorm(10, mean =mean(time_male))))
samp_mean_female<-replicate(1000, mean(rnorm(10, mean =mean(time_female))))
# Store means of male and female data in new variable
mM<-mean(samp_mean_male)
fM<-mean(samp_mean_female)
# calculate differnce
difference<-mM-fM

```

2. *Investigate whether the difference you observed in part 1 is significant, using procedures explored in the lecture. There are two groups in this investigation: male (`Melanoma$sex == 0`) and female (`Melanoma$sex == 1`). Randomly reassign the data in the `time` variable to the two groups, and compute the mean difference. Repeat 2000 times (this needs to be done relatively quickly; if it takes over a few minutes, I will dock points), and determine how frequently the difference in the mean survival time between men and women (that is, $\bar{x}_{\text{men}} - \bar{x}_{\text{women}}$ in the simulated data exceeds the same difference observed in the actual data. Does this analysis suggest the difference is due to "noise", or due to an actual difference in survival time between men and women?*

```{r, echo=FALSE}
# Seed reset; DO NOT CHANGE THE SEED!
set.seed(6222016)
```

```{r, tidy=TRUE}
# load MASS library
library(MASS)
# sample the male and female time data
m<-sample(time_male, 2000, replace=TRUE)
f<-sample(time_female, 2000, replace=TRUE)
# calculate the difference of this data 
diff<-mean(m < diff(f))
 diff
# diff value is .158
# this value is indeed significance and inicates tthat the difference in time between the sexes is not just noise
```

## Problem 4
*The data set `DDT` (**MASS**) contains measurements of the pesticide DDT in kale, in parts per million. Use bootstrapping to estimate a 95% confidence interval for the mean ppm of DDT in kale. Do the same with the standard deviation. Use 2000 replications each.*

```{r, echo=FALSE}
# Seed reset; DO NOT CHANGE THE SEED!
set.seed(6222016)
```

```{r, tidy=TRUE}
# Calculate mean
mean(DDT)
ddt_sim<-sample(DDT, size = length(DDT), replace = TRUE)
ddt_sim
mean(ddt_sim)
# bootstrap, while replicating 2000 times
mean_boot<-replicate(2000, {
  mean(sample(DDT, size = length(DDT), replace = TRUE))
})
# calculate .025 and .975 quantiles
quantile(mean_boot, c(.025, .975))
```